\documentclass[11pt]{article}

\usepackage{comun/pgm02sub} % NO TOCAR
\usepackage{graphicx} %Para incluir gráficas PS y PNG
\usepackage{url} % Para poner direcciones web


\title{Elvira: An Environment for Probabilistic Graphical Models}

\author{Elvira Consortium}

\summary{The Elvira system is a Java tool to construct
probabilistic decision support systems. Elvira
works with Bayesian networks and influence diagrams and it
can operate with discrete, continuous and temporal variables. It
has an easy to use Graphical User Interface (GUI). In addition you
can edit, easily, the ASCII format of the Elvira system to
introduce your models. It has methods for inference, learning,
abduction, fusion of knowledge and making decisions, so as some
tools to check the efficiency of the algorithms.} 

\pagestyle{empty}
\begin{document}
\maketitle

\section*{Elvira Consortium}
   
   {\small
   Members of the Elvira Consortium are: S. Acid, L.M.
   de Campos, A. Cano, J. Garc\'{\i}a, J. Huete, S. Moral
   (Universidad de Granada), J.A. G\'amez, J. Flores, C. Lacave,
   J.M. Puerta (Universidad de Castilla-La Mancha), L.D.
   Hern\'andez (Universidad de Murcia) I. Mart\'{\i}nez, F. Reche,
   C. Rodr\'{\i}guez, R. Rum\'{\i}, J. del Sagrado, A. Salmer\'on,
   F.J. Soler (Universidad de Almer\'{\i}a), M. G\'omez (Universidad de
   Ja\'en), F.J. D\'{\i}ez, S. Fern\'andez (UNED), I. Inza, P. Larra\~naga, B. Sierra
   (Universidad del Pa\'{\i}s Vasco), E. Sucar (Inst. Tecn. de
   Monterrey).
   }

\input introduction/introduction

\input format/format

\input gui/gui

\input inference/inference

\input id/id

\input learning/learning

\input future/future


\bibliographystyle{acl}
\bibliography{bib/Elvira-Presentation}

\newpage

$$
;
$$
\end{document}

